Trajectory Mapping Results

### 2021-03-24 scripts/analyze_trajectories.Rmd
analize traj rmd with tables

```{r parser, echo = FALSE}
library(tidyverse)
# Parser -----------------------------------------------------------------------
argsR <- commandArgs(TRUE)
args_list <- str_split(paste(argsR, collapse=""), pattern="--")
args_llist <- lapply(args_list, function(x) str_split(x, pattern="="))
argsDF <- as.data.frame(do.call("rbind", args_llist[[1]]))
args <- as.list(as.character(argsDF$V2))
names(args) <- argsDF$V1
args$burnin <- as.integer(args$burnin)
args$n <- as.integer(args$n)
title <- paste("Analysis", gsub("\\.", " ", gsub(".TL.traj", " particles", str_split(args$args, "/")[[1]][3])))
```

---
subtitle: "`r title`"
---

@@

```{r debugging}
# Debugging
args = list()
args$burnin = 0.1
args$demes = read.csv("./files/demes_China.csv", stringsAsFactors = FALSE)
args$args = "/Users/maceci/europe10last/210205_europe10.10000.TL.traj"
args$metadata =  "./results/europeC/final_metadata.tsv"
args$n = 100
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
demes <- demes %>% arrange(deme)
```

## Deme configuration
\tiny
```{r demeconfig}
# Load demes configuration and sequence metadata -------------------------------
demes <- data.frame(deme = NA, region = NA, country = NA, division = NA,
                    exclude_country = NA, exclude_division = NA,
                    min_date = NA, max_date = NA, seq_per_deme = NA) %>%
  bind_rows(bind_rows(lapply(yaml::yaml.load(string = gsub(",", ", ", gsub(":", ": ", args$demes))),
                          data.frame, stringsAsFactors = FALSE), .id = "deme")) %>%
  group_by(deme) %>%
  mutate(exclude_country = ifelse(is.na(exclude_country), NA, paste0(exclude_country, collapse = ","))) %>%
  distinct() %>%
  filter_all(any_vars(!is.na(.))) %>%
  mutate(deme = str_to_title(deme)) %>%
  mutate(deme = ifelse(deme %in% c("Othereuropean", "OtherEuropean"), "Other European", deme)) %>%
  arrange(deme)

@@

# Data wrangling ---------------------------------------------------------------
df_traj <- processEvents(df$events, demes, "China", mrs)
```

@@

first_ecdc <- case_data %>% 
  filter(source == "ecdc", total_confirmed != 0) %>% 
  group_by(deme) %>% 
  arrange(date) %>% 
  slice(1)

first_eui <- first_df1 %>% 
  ggplot(aes(date)) +
  geom_density(aes(y = ..count..), fill = "grey", alpha = .5, col = NA) + 
  geom_histogram(binwidth = 1, alpha = .5, fill = "#21908CFF", color = "#21908CFF") +
  geom_vline(aes(xintercept = median(first_df1$date), linetype = "first case median infered date")) +
  geom_vline(aes(xintercept = min((first_ecdc %>% filter(deme != "China"))$date), linetype = "first confirmed case to ECDC")) +
  ylab("epidemic trajectory count") +
  scale_x_date(limits = c(ymd("2019-12-15"), ymd("2020-02-01")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = c(1,2)) +
  theme(axis.title.x = element_blank())

first_euii <- first_df1 %>% 
  ggplot(aes(date)) +
  geom_density(fill = "grey", alpha = .5, col = NA) + 
  geom_histogram(binwidth = 1, alpha = .5, fill = "#21908CFF", color = "#21908CFF", aes( y = ..density..)) +
  geom_vline(aes(xintercept = median(first_df1$date), linetype = "first case median infered date")) +
  geom_vline(aes(xintercept = min((first_ecdc %>% filter(deme != "China"))$date), linetype = "first confirmed case to ECDC")) +
  ylab("probability density") +
  scale_y_continuous(position = "right") +
  scale_x_date(limits = c(ymd("2019-12-15"), ymd("2020-02-01")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = c(1,2)) +
  theme(axis.title.x = element_blank())

ggexport(first_eui,
         filename = paste0(args$output_figure, "03i.png"),
         width = 2100, height = 800, res = 300)
ggexport(first_euii,
         filename = paste0(args$output_figure, "03ii.png"),
         width = 2100, height = 800, res = 300)
```

## First imported cases into Europe, time and location
```{r first-eu-destsrc, fig.height=4}
# 2.2b First introduction into Europe, distribution density destination
facet_names <- demes$deme
names(facet_names) <- demes$deme
facet_names["OtherEuropean"] <- "Other EU"

first_eu1d <- ggplot(first_df1, aes(x = date, y = ..count../(args$n))) +
  geom_density_line(
    data = select(first_df1, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = deme), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE,
             labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First imported case into Europe, \ndestination") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

# 2.2b First 10 introduction into Europe, distribution density destination
m = 10
first_df10 <- df_traj %>%
  # 1. Prepare data
  filter(var == "IM", !is.na(partner),  value != 0) %>%
  group_by(traj) %>%
  arrange(date, partner) %>%
  slice(1:m)

first_eu10d <- ggplot(first_df10, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_df10, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = deme), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE, labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First 10 imported cases into Europe, \ndestination") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

first_eu10s <- ggplot(first_df10, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_df10, -partner), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = partner), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(partner), switch = "y", drop = FALSE, labeller = labeller(partner = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by source country") +
  xlab("date") +
  labs(title = "First 10 imported cases into Europe, \nsource") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

first_eu_ds <- ggarrange(first_eu1d, first_eu10d, first_eu10s, labels = chartr("123456789", "ABCDEFGHI", 1:3), font.label = list(size = 9), nrow = 1, ncol = 3, hjust = -6, vjust = 2)
first_eu_ds
ggexport(first_eu_ds,
         filename = paste0(args$output_figure, "04.png"),
         width = 2500, height = 1500, res = 300)

## First introductions into each country, source
```{r first-bycountry, fig.height=4}
# 2.2b First introduction into each country, distribution density destination
m = 1
first_country <- df_traj %>%
  mutate(partner = ifelse(var == "O", "China", partner)) %>%
  # 1. Prepare data
  filter(var == "IM" & deme != "China"| var == "O" & deme == "China", !is.na(partner), value != 0) %>%
  group_by(traj, deme) %>%
  arrange(date, partner) %>%
  slice(1:m)

first_10countryplot <- ggplot(first_10country, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_10country, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 3
  ) + 
  geom_density_line(aes(fill = partner), color = "transparent", bw = 3, alpha = 0.6) +
  geom_vline(data = first_ecdc, aes(xintercept = date, linetype = "first confirmed case to ECDC")) +
  #geom_vline(data = first_10country %>% group_by(deme) %>% summarise(median = median(date)), aes(xintercept = median, linetype = "first 10 imported cases median date")) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE, labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "source country", values = dcolors, na.value = "grey") +
  scale_x_date(#limits = c(ymd("2020-01-01"), ymd("2020-02-15")), 
    date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = 2) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First 10 imported cases into the country") +
  theme(#legend.position = "none",
        legend.key.size = unit(0.5, "cm"),
        axis.title.x=element_blank()) +
  guides(fill = guide_legend(nrow = 1))

ggexport(first_countryds,
         filename = paste0(args$output_figure, "05.png"),
         width = 2500, height = 1700, res = 300)
```

events_timingplot <- ggplot(df_eventsttiming) +
  geom_density_ridges(aes(x = date, y = factor(var, level =  c("OM", "IM", "D", "B")), fill = var), color = "transparent", quantile_lines = TRUE, quantiles = c(0.025, 0.975), bandwidth=2, alpha = 0.7, size = 0.2) + 
  geom_vline(data = first_ecdc, aes(xintercept = date, linetype = "first confirmed case to ECDC")) +
  scale_linetype_manual(values = 2, name = "") +
  facet_wrap(~deme) +
  scale_fill_viridis_d(guide = FALSE) +
  scale_y_discrete(breaks = c("B", "D", "IM", "OM"), labels = c("first local case", "first death/recovered/isolated", "first imported case", "first exported case")) + 
  ylab("") +
  xlab("")

events_timingplot
ggexport(events_timingplot,
         filename = paste0(args$output_figure, "13.png"),
         width = 2000, height = 1200, res = 300)

@@

## Summarised epidemic trajectories
```{r trajs ribbon, fig.height = 4}
# 1.1 Median and 95% CCI by deme in log scale
gribbon <- ggplot(traj_summary %>% filter(var == "active_pop")) +
  geom_ribbon(aes(date, ymin = l95_cumvalue, ymax = h95_cumvalue, fill = deme), alpha = 0.5) +
  geom_line(aes(date, median_cumvalue, colour = deme)) +
  ylab("active SARS-CoV-2 infections") +
  #xlab("Date") +
  #labs(title = "SARS-CoV-2 early epidemics in Europe and Hubei",
  #subtitle = "Case trajectories Median and 95% CCI by deme") +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  scale_x_date(limits = c(ymd("2019-10-01"), ymd(MRS)), date_breaks = "1 month", date_labels = "%b %d") +
  scale_fill_manual(name = "", values = dcolors) +
  scale_color_manual(name = "", values = dcolors) +
  theme(legend.position = c(0.2, 0.7),
        legend.box="horizontal",
        axis.title.x=element_blank())

@@

ggexport(gribbon,
         filename = paste0(args$output_figure, "01.png"),
         width = 1500, height = 1000, res = 300)
```

```{r trajs}
# 1.2 All trajectories single deme in log scale with reported cases date
trajs <- lapply(demes$deme, function(d) {
  df_traj_cases%>%
    # 1. Prepare data
    filter(deme == d, 
           var %in% c("D", "ecdc"), 
           is.na(partner)) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumvalue, group = traj,
                  color = var == "ecdc",
                  alpha = var == "ecdc",
                  linetype = var == "ecdc",
                  size = var == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(d, " inferred\n population trajectories"),
                                  paste0(ifelse(d == china_deme, "China", as.character(d)),
                                         " total reported cases")),
                       values = c(dcolors[[d]], dark_dcolors[[d]])) +
    scale_alpha_manual(name = "", values = c(0.1, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 5),
                          labels = c(paste0(d, " inferred\n population trajectories"),
                                     paste0(ifelse(d == china_deme, "China", as.character(d)),
                                            " total reported cases"))) +
    scale_size_manual(name = "", values = c(0.2, 1.5),
                      labels = c(paste0(d, " inferred\n population trajectories"),
                                 paste0(ifelse(d == china_deme, "China", as.character(d)),
                                        " total reported cases"))) +
    scale_x_date(limits = c(ymd("2019-10-01"), ymd(MRS)), date_breaks = "1 month",
                 date_labels = "%b %y") +
    scale_y_log10(limits= c(1,10^7.5), expand = c(0, 0),labels = trans_format("log10", math_format(10^.x))) +
    labs(title = d) +
    ylab(ifelse(d %in% c("China", "Germany", "Other European"), "SARS-CoV-2 cases", "")) +
    theme(axis.title.x=element_blank(),
          legend.position = "none",
          legend.box="horizontal",
          panel.grid.major.y =  element_line(colour = "grey80", size = 0.3, linetype = 2),
          panel.grid.minor.y =  element_line(colour = "grey80", size = 0.2,  linetype = 3),
          panel.background =  element_blank(),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 10)) 
})

names(trajs) <- demes$deme
ggexport(ggarrange(plotlist = trajs, labels = chartr("123456789", "ABCDEFGHI", 1:length(trajs)), nrow = 3, ncol = 2),
         filename = paste0(args$output_figure, "02.png"),
         width = 2300, height = 3000, res = 300)
```

@@

@@

ratio_plots <- lapply(demes$deme, function(d){
  hweek %>% left_join(inpop) %>% filter(deme == d) %>%
    ggplot() +
    geom_line(aes(year_week, R, group=traj, color = log(cumvalue)), alpha = 0.5) +
    geom_line(data = hsummary %>% filter(deme == d),aes(year_week, mR), size = 1) +
    geom_vline(aes(xintercept=ymd("2020-01-23")), linetype = "dashed") +
    geom_vline(aes(xintercept=ymd("2020-03-01")), linetype = "dashed") +
    #facet_wrap(~deme, scales = "free") +
    coord_cartesian(ylim = c(0,200)) +
    xlab("") +
    ylab(ifelse(d %in% c("China", "Italy"), "local/imported cases\n by week", "")) +
    labs(title = d) +
    #scale_colour_gradient(low = "yellow", high = "red", na.value = NA)
    scale_colour_viridis_c(name = "log cases", direction = -1) +
    theme(legend.position = c(0.2, 0.8),
          legend.key.size = unit(0.2, "cm"),
          title = element_text(size = 10))
})

@@

```{r circoplot}
plot_chord <- function(df, min_date, max_date) {
  # df <- df %>%
  #   filter(date > min_date, date <= max_date, var == "IM") %>%
  #   group_by(partner, deme, date) %>%
  #   # Take mean value over all trajectories
  #   summarise(Nmean = mean(value), .groups = "drop_last") %>%
  #   # Sum mean values of each day over all the time period
  #   summarise(N = sum(Nmean), .groups = "drop") %>%
  #   drop_na()

>>>>

```{r parser, echo = FALSE}
library(tidyverse)
# Parser -----------------------------------------------------------------------
# argsR <- commandArgs(TRUE)
# args_list <- str_split(paste(argsR, collapse=""), pattern="--")
# args_llist <- lapply(args_list, function(x) str_split(x, pattern="="))
# argsDF <- as.data.frame(do.call("rbind", args_llist[[1]]))
# args <- as.list(as.character(argsDF$V2))
# names(args) <- argsDF$V1
# args$burnin <- as.integer(args$burnin)
# args$n <- as.integer(args$n)
# title <- paste("Analysis", gsub("\\.", " ", gsub(".TL.traj", " particles", str_split(args$args, "/")[[1]][3])))
# ---
# subtitle: "`r title`"
# ---
```

@@

```{r debugging}
# Debugging
args = list()
args$burnin = 0.1
args$demes = read.csv("./files/demes_China.csv", stringsAsFactors = FALSE)
args$args = "/Users/maceci/europe10last/210205_europe10.10000.TL.traj"
args$metadata =  "./results/europeC/final_metadata.tsv"
args$n = 10
demes = args$demes
demes$exclude_country[5] = "France,Germany,Italy,Spain"
demes <- demes %>% arrange(deme)
```

## Deme configuration
\tiny
```{r demeconfig}
# Load demes configuration and sequence metadata -------------------------------
# demes <- data.frame(deme = NA, region = NA, country = NA, division = NA,
#                     exclude_country = NA, exclude_division = NA,
#                     min_date = NA, max_date = NA, seq_per_deme = NA) %>%
#   bind_rows(bind_rows(lapply(yaml::yaml.load(string = gsub(",", ", ", gsub(":", ": ", args$demes))),
#                           data.frame, stringsAsFactors = FALSE), .id = "deme")) %>%
#   group_by(deme) %>%
#   mutate(exclude_country = ifelse(is.na(exclude_country), NA, paste0(exclude_country, collapse = ","))) %>%
#   distinct() %>%
#   filter_all(any_vars(!is.na(.))) %>%
#   mutate(deme = str_to_title(deme)) %>%
#   mutate(deme = ifelse(deme %in% c("Othereuropean", "OtherEuropean"), "Other European", deme)) %>%
#   arrange(deme)

@@

# Data wrangling ---------------------------------------------------------------
df_traj <- processEvents(df$events, demes, "China", MRS)
```

@@

@@

first_ecdc <- case_data %>% 
  filter(source == "ecdc", total_confirmed != 0) %>% 
  group_by(deme) %>% 
  arrange(date) %>% 
  slice(1) %>%
  ungroup()

first_eui <- first_df1 %>% 
  ggplot(aes(date)) +
  geom_density(aes(y = ..count..), fill = "grey", alpha = .5, col = NA, bw = 2) + 
  geom_histogram(binwidth = 1, alpha = .5, fill = "#21908CFF", color = "#21908CFF") +
  geom_vline(aes(xintercept = median(first_df1$date), linetype = "first case median infered date")) +
  geom_vline(aes(xintercept = min((first_ecdc %>% filter(deme != "China"))$date), linetype = "first confirmed case to ECDC")) +
  ylab("epidemic trajectory count") +
  scale_x_date(limits = c(ymd("2019-12-15"), ymd("2020-02-01")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = c(1,2)) +
  theme(axis.title.x = element_blank())

first_euii <- first_df1 %>% 
  ggplot(aes(date)) +
  geom_density(fill = "grey", alpha = .5, col = NA, bw = 2) + 
  geom_histogram(binwidth = 1, alpha = .5, fill = "#21908CFF", color = "#21908CFF", aes( y = ..density..)) +
  geom_vline(aes(xintercept = median(first_df1$date), linetype = "first case median infered date")) +
  geom_vline(aes(xintercept = min((first_ecdc %>% filter(deme != "China"))$date), linetype = "first confirmed case to ECDC")) +
  ylab("probability density") +
  scale_y_continuous(position = "right") +
  scale_x_date(limits = c(ymd("2019-12-15"), ymd("2020-02-01")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = c(1,2)) +
  theme(axis.title.x = element_blank())

first_eui

ggexport(first_eui,
         filename = paste0(args$output_figure, "01i.png"),
         width = 2100, height = 800, res = 300)
ggexport(first_euii,
         filename = paste0(args$output_figure, "01ii.png"),
         width = 2100, height = 800, res = 300)
```

*Median date:* `r median(first_df1$date)`\
*95% CCI:* (`r quantile(first_df1$date, 0.025, type = 1)`, `r quantile(first_df1$date, 0.975, type = 1)`)\
*ECDC first confirmed case:* `r min((first_ecdc %>% filter(deme != "China"))$date)`\

## First imported cases to Europe, time and location
```{r first-eu-destsrc, fig.height=4}
# 2.2b First introduction to Europe, distribution density destination
facet_names <- demes$deme
names(facet_names) <- demes$deme
facet_names["OtherEuropean"] <- "Other EU"

first_eu1d <- ggplot(first_df1, aes(x = date, y = ..count../(args$n))) +
  geom_density_line(
    data = select(first_df1, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = deme), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE,
             labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First imported case to Europe, \ndestination") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

# 2.2b First 10 introduction to Europe, distribution density destination
m = 10
first_df10 <- df_traj %>%
  # 1. Prepare data
  filter(var == "IM", !is.na(partner),  value != 0) %>%
  group_by(traj) %>%
  arrange(date, partner) %>%
  slice(1:m)

first_eu10d <- ggplot(first_df10, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_df10, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = deme), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE, labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First 10 imported cases to Europe, \ndestination") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

first_eu10s <- ggplot(first_df10, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_df10, -partner), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 2) + 
  geom_density_line(aes(fill = partner), bw = 2, color = "transparent", alpha = 0.8) +
  facet_grid(rows = vars(partner), switch = "y", drop = FALSE, labeller = labeller(partner = facet_names)) +
  scale_fill_manual(name = "", values = dcolors, na.value = "grey") +
  scale_x_date(limits = c(ymd("2020-01-01"), ymd("2020-02-15")), date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  ylab("probability density by source country") +
  xlab("date") +
  labs(title = "First 10 imported cases to Europe, \nsource") +
  theme(legend.position = "none",
        axis.title.x=element_blank())

first_eu_ds <- ggarrange(first_eu1d, first_eu10d, first_eu10s, labels = chartr("123456789", "ABCDEFGHI", 1:3), font.label = list(size = 9), nrow = 1, ncol = 3, hjust = -6, vjust = 2)
first_eu_ds
ggexport(first_eu_ds,
         filename = paste0(args$output_figure, "02.png"),
         width = 2500, height = 1500, res = 300)

## First imported cases to Europe, time and location
```{r first-eu-table}
knitr::kable(first_df1 %>%
               ungroup %>%
               count(deme, name = "n_1d") %>%
               mutate(p_1d = n_1d/args$n) %>%
               left_join(first_df10 %>%
                           ungroup %>%
                           count(deme,  name = "n_10d") %>%
                           mutate(p_10d = n_10d/(10*args$n)) %>%
                           left_join(first_df10 %>%
                                       ungroup %>%
                                       select(-deme) %>% rename(deme = partner) %>%
                                       count(deme,  name = "n_10s") %>%
                                       mutate(p_10s = n_10s/(10*args$n)))))
```

## First introductions to each country, source
```{r first-bycountry, fig.height=4}
# 2.2b First introduction into each country, distribution density destination
m = 1
first_country <- df_traj %>%
  mutate(partner = ifelse(var == "O", "China", partner)) %>%
  # 1. Prepare data
  filter(var == "IM" & deme != "China"| var == "O" & deme == "China", !is.na(partner), value != 0) %>%
  group_by(traj, deme) %>%
  arrange(date, partner) %>%
  slice(1:m)

first_10countryplot <- ggplot(first_10country, aes(x = date, y = ..count../(m*args$n))) +
  geom_density_line(
    data = select(first_10country, -deme), aes(fill = NA),
    color = "transparent", alpha = 0.8, bw = 3
  ) + 
  geom_density_line(aes(fill = partner), color = "transparent", bw = 3, alpha = 0.6) +
  geom_vline(data = first_ecdc, aes(xintercept = date, linetype = "first confirmed case to ECDC")) +
  #geom_vline(data = first_10country %>% group_by(deme) %>% summarise(median = median(date)), aes(xintercept = median, linetype = "first 10 imported cases median date")) +
  facet_grid(rows = vars(deme), switch = "y", drop = FALSE, labeller = labeller(deme = facet_names)) +
  scale_fill_manual(name = "source country", values = dcolors, na.value = "grey") +
  scale_x_date(#limits = c(ymd("2020-01-01"), ymd("2020-02-15")), 
    date_breaks = "1 week", date_labels = "%b %d, %Y", expand = c(0,0)) +
  scale_linetype_manual(name = "", values = 2) +
  ylab("probability density by destination country") +
  xlab("date") +
  labs(title = "First 10 imported cases to the country") +
  theme(#legend.position = "none",
    legend.key.size = unit(0.5, "cm"),
    axis.title.x=element_blank()) +
  guides(fill = guide_legend(nrow = 1))

ggexport(first_countryds,
         filename = paste0(args$output_figure, "03.png"),
         width = 2500, height = 1700, res = 300)
```

## First introductions to each country, source

:::: {.column width="50%"}
\small
```{r first-country-table1}
knitr::kable(first_country %>%
               group_by(deme) %>%
               summarise(m = median(date),
                         l = quantile(date, 0.025, type = 1),
                         h = quantile(date, 0.925, type = 1)))
```
::::

:::: {.column width="50%"}
\small
```{r first-country-table2}

knitr::kable(first_country %>%
               ungroup %>%
               group_by(deme) %>%
               count(partner, name = "n_1s") %>%
               mutate(p_1s = n_1s/args$n))
```

## First introductions to each country, source

:::: {.column width="50%"}
\small
```{r first-country-table3}
knitr::kable((first_10country %>%
                ungroup %>%
                group_by(deme) %>%
                count(partner, name = "n_10s") %>%
                mutate(p_10s = n_10s/(10 * args$n)))[1:15,])
```
::::

:::: {.column width="50%"}
\small
```{r first-country-table4}
knitr::kable((first_10country %>%
                ungroup %>%
                group_by(deme) %>%
                count(partner, name = "n_10s") %>%
                mutate(p_10s = n_10s/(10 * args$n)))[16:30,])
```

events_timingplot <- ggplot(df_eventsttiming) +
  geom_density_ridges(aes(x = date, y = factor(var, level =  c("OM", "IM", "D", "B")), fill = var), color = "transparent", quantile_lines = TRUE, quantiles = c(0.025, 0.975), bandwidth=2, alpha = 0.7, size = 0.2) + 
  geom_vline(data = first_ecdc, aes(xintercept = date, linetype = "first confirmed case to ECDC")) +
  scale_linetype_manual(values = 2, name = "") +
  facet_wrap(~deme) +
  scale_fill_viridis_d(guide = FALSE) +
  scale_y_discrete(breaks = c("B", "D", "IM", "OM"), labels = c("first local case", "first death/recovered/isolated", "first imported case", "first exported case")) + 
  scale_x_date(#limits = c(ymd("2020-01-01"), ymd("2020-02-15")), 
    date_breaks = "1 month", date_labels = "%b, %Y", expand = c(0,0)) +
  ylab("") +
  xlab("")

events_timingplot
ggexport(events_timingplot,
         filename = paste0(args$output_figure, "04.png"),
         width = 2000, height = 1200, res = 300)

:::: {.column width="50%"}
\small
```{r events-timing-table1}
knitr::kable((df_eventsttiming %>%
                group_by(deme, var) %>%
                summarise(m = median(date),
                          l = quantile(date, 0.025, type = 1),
                          h = quantile(date, 0.925, type = 1)))[1:12,])
```
::::

:::: {.column width="50%"}
\small
```{r events-timing-table2}
knitr::kable((df_eventsttiming %>%
                group_by(deme, var) %>%
                summarise(m = median(date),
                          l = quantile(date, 0.025, type = 1),
                          h = quantile(date, 0.925, type = 1)))[12:24,])
```

## Summarised epidemic trajectories

```{r trajs ribbon, fig.height = 4}
# 1.1 Median and 95% CCI by deme in log scale
gribbon <- ggplot(traj_summary %>% filter(var == "active_pop")) +
  geom_ribbon(aes(date, ymin = l95_cumvalue, ymax = h95_cumvalue, fill = deme), alpha = 0.5) +
  geom_line(aes(date, median_cumvalue, colour = deme)) +
  ylab("active SARS-CoV-2 infections") +
  #xlab("Date") +
  #labs(title = "SARS-CoV-2 early epidemics in Europe and Hubei",
  #subtitle = "Case trajectories Median and 95% CCI by deme") +
  scale_y_log10(labels = trans_format("log10", math_format(10^.x))) +
  scale_x_date(limits = c(ymd("2019-10-01"), ymd(MRS)), date_breaks = "1 month", date_labels = "%b, %Y", expand = c(0,0)) +
  scale_fill_manual(name = "", values = dcolors) +
  scale_color_manual(name = "", values = dcolors) +
  theme(legend.position = c(0.2, 0.7),
        legend.box="horizontal",
        axis.title.x=element_blank())

@@

ggexport(gribbon,
         filename = paste0(args$output_figure, "05.png"),
         width = 1500, height = 1000, res = 300)
```

```{r trajs}
# 1.2 All trajectories single deme in log scale with reported cases date
trajs <- lapply(demes$deme, function(d) {
  df_traj_cases%>%
    # 1. Prepare data
    filter(deme == d, 
           var %in% c("D", "ecdc"), 
           is.na(partner)) %>%
    # 2. Plot
    ggplot() +
    geom_line(aes(date, cumvalue, group = traj,
                  color = var == "ecdc",
                  alpha = var == "ecdc",
                  linetype = var == "ecdc",
                  size = var == "ecdc")) +
    scale_color_manual(name = "",
                       labels = c(paste0(d, " inferred\n population trajectories"),
                                  paste0(ifelse(d == china_deme, "China", as.character(d)),
                                         " total reported cases")),
                       values = c(dcolors[[d]], dark_dcolors[[d]])) +
    scale_alpha_manual(name = "", values = c(0.1, 1), guide = FALSE) +
    scale_linetype_manual(name = "", values = c(1, 5),
                          labels = c(paste0(d, " inferred\n population trajectories"),
                                     paste0(ifelse(d == china_deme, "China", as.character(d)),
                                            " total reported cases"))) +
    scale_size_manual(name = "", values = c(0.2, 1.5),
                      labels = c(paste0(d, " inferred\n population trajectories"),
                                 paste0(ifelse(d == china_deme, "China", as.character(d)),
                                        " total reported cases"))) +
    scale_x_date(limits = c(ymd("2019-10-01"), ymd(MRS)), date_breaks = "1 month", date_labels = "%b, %Y", expand = c(0,0)) +
    scale_y_log10(limits= c(1,10^7.5), expand = c(0, 0),labels = trans_format("log10", math_format(10^.x))) +
    labs(title = d) +
    ylab(ifelse(d %in% c("China", "Germany", "Other European"), "SARS-CoV-2 cases", "")) +
    theme(axis.title.x=element_blank(),
          legend.position = "none",
          legend.box="horizontal",
          panel.grid.major.y =  element_line(colour = "grey80", size = 0.3, linetype = 2),
          panel.grid.minor.y =  element_line(colour = "grey80", size = 0.2,  linetype = 3),
          panel.background =  element_blank(),
          axis.text.x = element_text(size = 8),
          axis.text.y = element_text(size = 10)) 
})

names(trajs) <- demes$deme
ggexport(ggarrange(plotlist = trajs, labels = chartr("123456789", "ABCDEFGHI", 1:length(trajs)), nrow = 3, ncol = 2),
         filename = paste0(args$output_figure, "06.png"),
         width = 2300, height = 3000, res = 300)
```

@@

report_plot
ggexport(report_plot,
         filename = paste0(args$output_figure, "07.png"),
         width = 2000, height = 1200, res = 300)
```

\small
```{r report-prop-table}
knitr::kable(traj_summary %>%
               filter(var %in% c("D")) %>%
               left_join(case_data %>% filter(source == "ecdc")) %>%
               select(deme, date, l95_cumvalue, h95_cumvalue, median_cumvalue, total_confirmed) %>%
               group_by(deme) %>%
               filter(date == max(date)) %>%
               mutate(median_report = total_confirmed/median_cumvalue,
                      h95_report = total_confirmed/l95_cumvalue,
                      l95_report = total_confirmed/h95_cumvalue))
```

@@

ratio_plots <- lapply(demes$deme, function(d){
  hweek %>% left_join(inpop) %>% filter(deme == d) %>%
    ggplot() +
    geom_line(aes(year_week, R, group=traj, color = log(cumvalue)), alpha = 0.5) +
    geom_line(data = hsummary %>% filter(deme == d),aes(year_week, mR), size = 1) +
    geom_vline(aes(xintercept=ymd("2020-01-23")), linetype = "dashed") +
    geom_vline(aes(xintercept=ymd("2020-03-01")), linetype = "dashed") +
    #facet_wrap(~deme, scales = "free") +
    coord_cartesian(ylim = c(0,200)) +
    xlab("") +
    ylab(ifelse(d %in% c("China", "Italy"), "ratio local/imported\nnew cases by week", "")) +
    labs(title = d) +
    #scale_colour_gradient(low = "yellow", high = "red", na.value = NA)
    scale_x_date(date_labels = "%b %d, %Y") +
    scale_colour_viridis_c(name = "log cases", direction = -1) +
    theme(legend.position = c(0.2, 0.8),
          legend.key.size = unit(0.2, "cm"),
          title = element_text(size = 10))
})

@@

```{r bvsm-table, results = 'asis'}
for (d in demes$deme) {
  cat("\n## Local transmission vs imported cases - ", d, "\n")
  knitr::kable(hsummary %>% filter(deme == d))
  cat("\n")
}
```
